\mytitle

### 2023-05-13 chapters/intro.tex
Emina intro edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Like software development for any domain, systems software development requires
three fundamental tasks:
\textbf{specifying} correct behavior at a high level,
\textbf{implementing} the system at a low level,
and \textbf{verifying} that the implementation correctly adheres to the specification.
Where systems software distinguishes itself is in the level of abstraction
at which the implementation operates.
Specifically, many systems must operate between layers of abstraction,
such as between source and target languages or between the application and
hardware layers of the operating system.
This means that manually implementing and verifying systems software requires
a high amount of effort from the developer, who must reason about the multiple abstraction
layers in addition to the execution of their implemented system.
Additionally, stringent requirements on performance for systems
further increase the difficulty of implementation and verification.

A well-studied technique for reducing the burden of both implementation
and verification from software designers is \textit{program synthesis}.
Program synthesis is an automated programming technique
that takes advantage of overlapping semantics between the three tasks described above
in order to automate software implementation and verification. 
To do so, synthesis tools take as input a specification and output both an implementation
as well as a certificate verifying that the implementation correctly behaves
according to the input specification.

@@

To overcome these unique challenges, this dissertation presents a variety of novel
\textit{system decomposition} and \textit{metaprogramming abstraction} techniques.
%In this work, I identify two key observations that enable the creation of synthesis tools
\textbf{Program decomposition} allows synthesis tools to generate smaller target programs.
By considering systems as a conjunction of smaller pieces,
synthesis tools can search for candidate programs that implement a single piece.
The resulting synthesized program can be combined with either user-written code or
other synthesized system pieces, effectively reducing the developer burden.
% JitSynth: break up into minicompilers (TODO)
% DepSynth: consider storage mechanism separately from crash consistency mechanism (TODO)
Second, system-specific \textbf{metaprogram abstractions} enable synthesis tools to reason 
about candidate programs that themselves take DSL programs as input.
By reasoning about these simpler abstractions rather than the complex systems directly,
synthesis tools are able to more efficiently search over candidate programs.
% JitSynth: design metasketch to describe common minicompiler patterns (TODO)
% DepSynth: design rules language to describe any runtime dependencies (TODO)
In the next two sections, I describe how these observations enabled the creation of
synthesis tools for both in-kernel JIT compliers and for crash consistent storage systems.

@@

\section{Synthesis for In-Kernel JIT Compilers}
Modern operating systems have become increasingly extensible and customizable.
One way these systems enable customization is by
allowing users to submit custom code for kernel execution
written in an \textit{in-kernel DSL}.
Initially designed for network packet filtering, 
in-kernel DSLs like eBPF \cite{fleming:ebpf} enable kernel extentions
such as performance monitoring, load balancing, and intrusion detection
\cite{engler:vcode,fleming:ebpf,mccanne:bpf}.
For performance reasons, operating systems use just-in-time (JIT) compilers
to execute in-kernel DSL programs.
However,
% as mentioned previously, % TODO?
these JIT compilers can be the source of major kernel bugs
\cite{gpz:1454,paul:cve-2020-8835,blazakis:jit-spraying,kocher:spectre}.
Moreover, writing JIT compilers can pose a substantial effort to developers,
requiring on the order of thousands of lines of code.

In \cref{c:jitsynth}, I present \jitsynth, a program synthesis tool that
automatically generates JIT compilers for in-kernel DSLs
given a specification of the source and target language.
Existing work has demonstrated how the correctness of JIT compilers
can be proven manually\cite{wang:jitk,sobel:ejitk},
but the effort of such proofs is highly non-trivial.
By synthesizing JIT compilers, developers both avoid the burden of implementing
such systems \textit{and} can guarantee correctness
without requiring intensive proofs.

@@

\section{Automatic Crash Consistency with Synthesis}
Storage systems face an increasing demand for reliability.
From Linux file systems to large cloud-based key-value stores,
storage systems must maintain the integrity of their data
even in the presence of system crashes.
This property of storage systems is known as \textit{crash consistency}.
Guaranteeing crash consistency is a highly non-trivial task for developers,
as it requires maintaining complex invariants that depend greatly on
storage system implementation details.
Moreover, failing to guarantee crash consistency may result in severe impacts,
including costly slowdowns and massive data losses. % TODO make more concrete
% TODO finish: Highlight what can go wrong if crash consistency is done incorrectly
% TODO
% DepSynth: consider storage mechanism separately from crash consistency mechanism (TODO)
% DepSynth: design rules language to describe any runtime dependencies (TODO)

@@

\cref{c:depsynth} presents \depsynth, a tool for automating
crash consistency in storage systems. In \cref{c:depsynth},
we introduce a new dependency rule language that describes rules
which specify required write-before relationships for a storage system
at runtime. This abstraction over write-before relationships, along
with a new per-litmus-test search, allows \depsynth to synthesize
correct dependency rules for a production system.
Moreover, these generated rules have nearly the same effect on performance
as dependencies hand-written by experts.

Both of these chapters demonstrate
how program synthesis can aid developers in building
correct and performant systems.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Like software development for any domain, systems software development involves
three fundamental tasks:
\textbf{specifying} correct behavior at a high level,
\textbf{implementing} the system at a low level,
and \textbf{verifying} that the implementation correctly adheres to the specification.
Where systems software distinguishes itself is in the level of abstraction
at which the implementation operates.
Specifically, many systems must operate between layers of abstraction,
such as between source and target languages or between the application and
hardware layers of the operating system.
This means that manually implementing and verifying systems software requires
a high amount of effort from the developer, who must reason about the multiple abstraction
layers in addition to the execution of their implemented system.
Additionally, stringent requirements on performance of systems software
further increase the difficulty of implementation and verification.

A well-studied technique for reducing the burden of software implementation
and verification is \textit{program synthesis}.
Program synthesis is an automated programming technique
that takes advantage of overlapping semantics between the three tasks described above
in order to automate software implementation and verification. 
To do so, synthesis tools take as input a specification and output both an implementation
as well as a certificate verifying that the implementation correctly behaves
according to the input specification.

@@

To overcome these challenges, this dissertation presents novel
\textit{system decomposition} and \textit{metaprogramming abstraction} techniques.
%In this work, I identify two key observations that enable the creation of synthesis tools
\textbf{System decomposition} allows synthesis tools to generate smaller target programs.
By considering systems as a conjunction of smaller pieces,
synthesis tools can search for candidate programs that implement a single piece.
The resulting synthesized program can be combined with either user-written code or
other synthesized system pieces, effectively reducing the developer burden.
% JitSynth: break up into minicompilers (TODO)
% DepSynth: consider storage mechanism separately from crash consistency mechanism (TODO)
Second, system-specific \textbf{metaprogram abstractions} enable synthesis tools to reason 
about candidate programs that themselves take DSL programs as input.
By reasoning about these simpler abstractions rather than the complex systems directly,
synthesis tools are able to more efficiently search over candidate programs.
% JitSynth: design metasketch to describe common minicompiler patterns (TODO)
% DepSynth: design rules language to describe any runtime dependencies (TODO)
% TODO takeout the I here?
In the next two sections, I describe how these observations enabled the creation of
synthesis tools for both in-kernel JIT compliers and for crash consistent storage systems.

@@

\section{Synthesis for In-Kernel JIT Compilers}
Modern operating systems have become increasingly extensible and customizable.
One way these systems enable customization is by
allowing users to submit custom code for kernel execution
written in an \textit{in-kernel DSL}.
Initially designed for network packet filtering, 
in-kernel DSLs like eBPF \cite{fleming:ebpf} enable kernel extensions
such as performance monitoring, load balancing, and intrusion detection
\cite{engler:vcode,fleming:ebpf,mccanne:bpf}.
For performance reasons, operating systems use just-in-time (JIT) compilers
to execute in-kernel DSL programs.
However,
% as mentioned previously, % TODO?
these JIT compilers can be the source of major kernel bugs
\cite{gpz:1454,paul:cve-2020-8835,blazakis:jit-spraying,kocher:spectre}.
Moreover, writing JIT compilers can pose a substantial effort to developers,
requiring on the order of thousands of lines of code.

\cref{c:jitsynth} presents \jitsynth, a program synthesis tool that
automatically generates JIT compilers for in-kernel DSLs
given a specification of the source and target language.
Existing work has demonstrated how the correctness of JIT compilers
can be proven manually\cite{wang:jitk,sobel:ejitk},
but the effort of such proofs is non-trivial.
By synthesizing JIT compilers, developers both avoid the burden of implementing
such systems \textit{and} can guarantee correctness
without requiring intensive proofs.

@@

\section{Automatic Crash Consistency with Synthesis}
Storage systems face an increasing demand for reliability.
From Linux file systems to large cloud-based key-value stores,
storage systems must maintain the integrity of their data
even in the presence of system crashes.
This property of storage systems is known as \textit{crash consistency}.
Guaranteeing crash consistency is a challenging task for developers,
as it requires maintaining complex invariants that depend on
storage system implementation details.
Moreover, failing to guarantee crash consistency may result in severe impacts,
including costly slowdowns and massive data losses. % TODO make more concrete
% TODO finish: Highlight what can go wrong if crash consistency is done incorrectly
% TODO
% DepSynth: consider storage mechanism separately from crash consistency mechanism (TODO)
% DepSynth: design rules language to describe any runtime dependencies (TODO)

@@

\cref{c:depsynth} presents \depsynth, a tool for automating
crash consistency in storage systems. In \cref{c:depsynth},
we introduce a new dependency rule language that describes rules
which specify required write-before relationships for a storage system
at runtime. This abstraction over write-before relationships, along
with a new per-litmus-test search, allows \depsynth to synthesize
correct dependency rules for a production system.
Moreover, the generated rules have nearly the same effect on performance
as dependencies hand-written by experts.
These contributions demonstrate
how program synthesis can aid developers in building
correct and performant systems.
